\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-16 k-planar.tex
Added references, a comment about 1_planar_is_4_framed

This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs. \todo{Move this comment to the section on $1$-planar graphs?} For simple 1-planar graphs it is essential that we allow frames with parallel edges. \note{DW}{I feel that at this point we need to justify the introduction of multigraphs.}

@@

\todo[inline]{Alternative to the final three paragraphs in the preceding proof:

To see that $H$ is planar, consider a plane multigraph $G''$ obtained by adding stars in some of the faces of $G'$, as follows:  When creating the part $S$, add a vertex $v_\tau$ to the interior of $\tau$ that is adjacent to each vertex of $\tau$ and place $v_\tau$ in $S$.  The graph $G''\supseteq G'$ obtained this way is clearly planar.  Furthermore, for each part $P\in\mathcal{P}$, $G''[P]$ is connected.  Since planarity is preserved under edge contraction, it follows that $G''/\mathcal{P}$ is planar.  Now argue that $G/\mathcal{P}=G''/\mathcal{P}$\ldots
}

\note{DW}{The proof via 3-simplicity is shorter, so just keep it as is.}

@@

\begin{proof}[Proof of \cref{d_framed_product_stucture}]
	Let $G$ be a $d$-framed multigraph with frame $G'$ having outer face $F_0$.	Let $T$ be a BFS forest of $G'$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.  Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.  Now apply \cref{induction} on $G$, $G'$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
    This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.  Since any graph having a $3$-simple tree decomposition is planar \cite{X}, this implies that $H$ is planar.

@@

\note{DW}{Say something about how \cref{1_planar_is_4_framed} is better than similar results in the literature.}
\note{PM}{I don't know what to say without explaining that $4$-framed (simple) graphs only contain $3$-connected $1$-planar graphs.  Vida's example shows that there is a $1$-plane embedding of a graph that can't be augmented into a $4$-framed graph, but I haven't checked if it can't be drawn some better way. And I also don't know if the Bekos claim that $8$-framed (smple) graphs contain all $1$-planar graphs is proven anywhere and, if so, is it tight?}

@@

\begin{proof}
Let $G$ be a $1$-plane graph. We may assume that no two edges incident to a common vertex of $G$ cross, since such a crossing can be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}.
  % We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.  This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.

>>>>

This result is used below to obtain product structure theorems for 1-planar multigraphs and for map graphs. 
% \todo{Move this comment to the section on $1$-planar graphs?} 
For simple 1-planar graphs it is essential that we allow frames with parallel edges. 
% \note{DW}{I feel that at this point we need to justify the introduction of multigraphs.}

@@

% \todo[inline]{Alternative to the final three paragraphs in the preceding proof:
% 
% To see that $H$ is planar, consider a plane multigraph $G''$ obtained by adding stars in some of the faces of $G'$, as follows:  When creating the part $S$, add a vertex $v_\tau$ to the interior of $\tau$ that is adjacent to each vertex of $\tau$ and place $v_\tau$ in $S$.  The graph $G''\supseteq G'$ obtained this way is clearly planar.  Furthermore, for each part $P\in\mathcal{P}$, $G''[P]$ is connected.  Since planarity is preserved under edge contraction, it follows that $G''/\mathcal{P}$ is planar.  Now argue that $G/\mathcal{P}=G''/\mathcal{P}$\ldots
% }
% 
% \note{DW}{The proof via 3-simplicity is shorter, so just keep it as is.}

@@

\begin{proof}[Proof of \cref{d_framed_product_stucture}]
	Let $G$ be a $d$-framed multigraph with frame $G'$ having outer face $F_0$.	Let $T$ be a BFS forest of $G'$ rooted at $V(F_0)$ and let $\langle L_0,L_1,\ldots\rangle$ be the resulting BFS layering.  Let $P_1:= V(F_0)$, let $Y_1$ be any subset of $\min\{|P_1|,3\}$ vertices in $P_1$ and let $X_1:=P_1\setminus Y_1$.  Now apply \cref{induction} on $G$, $G'$, $T$, $F=F_0$, $k=1$, and $P_1=Y_1\cup X_1$.
    This gives an $H$-partition $\mathcal{P}:=\{S_x:x\in V(H)\}$ of $G$ in which $H$ has a $3$-simple tree decomposition.  Since any graph having a $3$-simple tree decomposition is planar \cite{knauer.ueckerdt:simple,kratochvil.vaner:note,elmallah.colbourn:on}, this implies that $H$ is planar.

@@

\note{DW}{Say something about how \cref{1_planar_is_4_framed} is better than similar results in the literature.}

\note{PM}{See my attempt below.}
% \note{PM}{I don't know what to say without explaining that $4$-framed (simple) graphs only contain $3$-connected $1$-planar graphs.  Vida's example shows that there is a $1$-plane embedding of a graph that can't be augmented into a $4$-framed graph, but I haven't checked if it can't be drawn some better way. And I also don't know if the Bekos claim that $8$-framed (smple) graphs contain all $1$-planar graphs is proven anywhere and, if so, is it tight?}

@@

Before proving \cref{1_planar_is_4_framed}, we remark that the lemma seems not to be true if we replace `multigraph' with `graph'.  We are only aware of two previous results on the relationship between $1$-planar graphs and $d$-framed graphs \cite{BDGGMR}:
\begin{inparaenum}[(i)]
  \item Every $3$-connected $1$-planar graph is contained in some $4$-framed graph.
  \item Every $1$-planar graph is contained in some $8$-framed graph.
\end{inparaenum}
Neither of these results allows a proof of \cref{1-planar} without sacrificing generality or accepting a larger constant.

\begin{proof}[Proof of \cref{1_planar_is_4_framed}]
Let $G$ be a $1$-plane graph. We may assume that no two edges incident to a common vertex of $G$ cross, since such a crossing can be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}.
  % We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.
We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add the edge $vw$, embedded in the face $F$.  This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.
